A Bayesian Nonparametric Approach to Neural Image Segmentation

### 2013-01-31 latex/submission.tex
model in paper is no longer a lie
graphical model set

\documentclass[draft,english]{article}
\usepackage{fullpage}
\usepackage{amsmath}

@@

\section{Model and Methods}

@@

We present our model in section~ref{sec:model}. The rest of the subsections explain how we overcome these difficulties. First, the number of cells must be inferred from the image itself. To do so, we employ a Mixture Dirichlet process, discussed in section~\ref{sec:mdp}. Further, the visual complexity and noise of electron micrograph images do not yield the clear separations required for assumptions 2 and 3. We fix this with Radon-like features to separate cell interiors and boundaries in section~\ref{sec:radon}, and constrain cluster assignments with a geodesic distance that penalizes boundary crossings in section~\ref{sec:geodesic}.

\subsection{Model}
\ref{sec:model}
In summary, pixel $i$ has feature vector $\+{x}_n = (x_n, y_n, r_n, f_n, \+{d}_n)$ where $x_n, y_n$ represent the pixel's spatial coordinates, $r_n \in [0,1]$ is the raw pixel intensity, $f_n \in [0,1]$ is the average response of the interior Radon-like feature, and $\+{d}_n \in \+{R}^N$ is a vector of geodesic distances  to every other pixel. The first four components, $\+{x}_{n,1:4}$ obey a multivariate normal distribution while the  Putting these components together yields the model in Figure~\ref{fig:model}. The generative story is:
\begin{itemize}
\item $\+\pi \sim \mbox{DirichletProcess}(\alpha)$
\item $k | \+\pi \sim \mbox{Discrete}(\+\pi)$
\item If $k$ is a new (empty) cluster:
  \begin{itemize}
  \item $\Lambda_k | T, \nu \sim \mbox{Wishart}(\Lambda | T, \nu)$ (in the notation of \cite{Murphy2003})
  \item $\+\mu_k | \kappa , \Lambda_k \sim \mathcal{N}\left( \+\mu_k | \+\mu_0 , (\kappa\Lambda)^{-1} \right)$
  \item $\beta_k \sim \mbox{Gamma}(\beta_k | \alpha_0, \beta_0)$
  \end{itemize}
\item $\+x | \+\mu_k , \Lambda_k, \beta_k \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \mbox{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \mbox{rate} = \beta_k \right)$ (in the notation of section~\ref{sec:geodesic})
\end{itemize}

@@

\subsection{Geodesic Distance}
\label{sec:geodesic}
Figure~\ref{fig:radon_bd} illustrates that cells need not be elliptical, or even convex. All that can be said is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response. Therefore, shape constraints based on Euclidean distances are not justified. Instead, we define a \emph{geodesic distance}. Construct a graph whose vertices are the pixels and undirected edge for adjacent pixels ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$). Each edge has a weight $w(i,j) = \left| b_i - b_j \right|$ where $b_k$ is the boundary Radon-like response at pixel $k$. We define $d(\+{x}_i, \+{x}_j)$ to be the total weight of the shortest path between pixels $i$ and $j$, which is computed with Johnson's algorithm. In practice, the $O(N^2)$ storage is prohibitive, so we instead we downsample the image into SLIC superpixels \cite{Achanta2012} and compute superpixel-based distances.Address practical concerns. Further, because individual cells are tiny relative to an entire slice, only local distances are needed, because far points must be outside the current cell. All paths crossing a cell boundary must traverse pixels with low, high, and again low boundary feature responses. Thus, the shortest path wil also be one with a high weight. So if pixels $i$ and $j$ do not belong to the same cell, defined by the region enclosed by a loop of pixels with high boundary feature response, then $d(\+{x}_i, \+{x}_j)$ is high.

@@

We model a two-dimensional image as an infinite semi-Gaussian mixture model. This extends the classic Gaussian mixture model (GMM) with (a) a variable number of classes and (b) additional naive-Bayes features associated with each class.

\begin{enumerate}
\item $\phi_{1\cdot} \in [0,1]$: pixel intensity 
\item $\phi_{2\cdot} \in [0,1]$: average interior Radon-like feature response
\item $\phi_{3\cdot} \in [0,1]$: geodesic distance  
\item 
\end{enumerate}
 Because cell interiors contain diverse structures, intensities values within a cell vary widely, so relying on $\phi_{1\cdot}$ alone oversegments. To reduce oversegmentation, we compute a cel boundary Radon-like feature for $\phi_{1\cdot}$, which homogenizes pixel intensities within a given cell~\cite{Kumar2010}. One computes Radon-like functions by taking line integrals of a kernel function along various scanning orientations, broken piecewise at image discontinuities. By varying the kernel, different structures and boundaries are enhanced. We follow Kumar's approach of averaging the integrals along each scan orientation to obtain a scalar response at each pixel. Finally, recall that our background class ($k = 1$) contains all pixels belonging on the cell membrane or in intracellular space. We penalize points

%, the cell Background Radon-like feature response in $\phi_{2\cdot}$, and the length of the shortest cluster path from $(

\section{Results}

>>>>

\documentclass[draft,english]{article}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{tkz-base}
\usetikzlibrary{shapes,decorations,shadows}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{decorations.shapes}
\usetikzlibrary{fadings}
\usetikzlibrary{patterns}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.footprints}
\usetikzlibrary{decorations.fractals}
\usetikzlibrary{shapes.gates.logic.IEC}
\usetikzlibrary{shapes.gates.logic.US}
\usetikzlibrary{fit,chains}
\usetikzlibrary{positioning}
\usepgflibrary{shapes}
\usetikzlibrary{scopes}
\usetikzlibrary{arrows}
\usetikzlibrary{backgrounds}
\tikzset{latent/.style={circle,fill=white,draw=black,inner sep=1pt, 
minimum size=20pt, font=\fontsize{10}{10}\selectfont},
obs/.style={latent,fill=gray!33},
const/.style={rectangle, inner sep=1pt},
factor/.style={rectangle, fill=black,minimum size=5pt, inner sep=1pt},
>={triangle 45}}

\pgfdeclarelayer{b}
\pgfdeclarelayer{f}
\pgfsetlayers{b,main,f}

\newcommand{\plate}[4]{
\begin{pgfonlayer}{b}
\node (invis#1) [draw, color=white, inner sep=6pt,rectangle,fit=#2] {};
\end{pgfonlayer}\begin{pgfonlayer}{f}
\node (capt#1) [ below left=0 pt of invis#1.south east, xshift=1pt,yshift=1pt] {\footnotesize{#3}};
\node (#1) [draw,inner sep=1pt, rectangle,fit=(invis#1) (capt#1),#4] {};
\end{pgfonlayer}
}

\newcommand{\shiftedplate}[5]{
\begin{pgfonlayer}{b}
\node (invis#1) [draw, color=white, inner sep=0 pt,rectangle,fit=#2] {};
\end{pgfonlayer}\begin{pgfonlayer}{f}
\node (capt#1) [#5, xshift=2pt] {\footnotesize{#3}};
\node (#1) [draw,inner sep=2pt, rectangle,fit=(invis#1) (capt#1),#4] {};
\end{pgfonlayer}
}

@@

\section{Model and Features}

@@

\begin{figure}
\centering

\begin{tikzpicture}
\matrix[row sep=6mm, column sep=8mm, matrix anchor=mid] (mdp) {
  \node (alpha) [const] {$\alpha$}; & & \\
  \node (pi) [latent] {$\+\pi$}; & \node (beta) [latent] {$\beta_k$}; & \node (alphabeta) [const] {$\alpha_0, \beta_0$}; \\
  \node (z) [latent] {$z_n$}; & \node(Lambda) [latent] {$\Lambda_k$}; & \node(Lambdanu) [const] {$\Lambda_0, \nu$}; \\
  \node (x) [obs] {$\+x_n$}; & \node(mu) [latent] {$\+\mu_k$}; & \node(mukappa) [const] {$\+\mu_0, \kappa$}; \\
};

\draw [->] (alpha) -- (pi);
\draw [->] (pi) -- (z);
\draw [->] (z) -- (x);
\draw [->] (alphabeta) -- (beta);
\draw [->] (Lambdanu) -- (Lambda);
\draw [->] (mukappa) -- (mu);
\draw [->] (Lambda) -- (mu);
\draw [->] (beta) -- (x);
\draw [->] (Lambda) -- (x);
\draw [->] (mu) -- (x);

\plate{Nplate}{(z)(x)}{N}{}
\plate{Kplate}{(beta)(Lambda)(mu)}{K}{}

\end{tikzpicture}
\caption{The graphical model} \label{fig:model}
\end{figure}

\label{sec:model}
We present our model in Figure~\ref{fig:model}. Below, we explain our feature representation and how we overcome the aforementioned difficulties. First, the number of cells must be inferred from the image itself. To do so, we employ a Mixture Dirichlet process, discussed in section~\ref{sec:mdp}. Further, the visual complexity and noise of electron micrograph images do not yield the clear separations required for assumptions 2 and 3. We fix this with Radon-like features to separate cell interiors and boundaries in section~\ref{sec:radon}, and constrain cluster assignments with a geodesic distance that penalizes boundary crossings in section~\ref{sec:geodesic}.

Pixel $i$ has feature vector $\+{x}_n = (x_n, y_n, r_n, f_n, \+{d}_n)$ where $x_n, y_n$ represent the pixel's spatial coordinates, $r_n \in [0,1]$ is the raw pixel intensity, $f_n \in [0,1]$ is the average response of the interior Radon-like feature, and $\+{d}_n \in \+{R}^N$ is a vector of geodesic distances to every other pixel. If position $\+p \in \+R^2$, when discretized, is the location of pixel $n$ and the discrete $\+q$ is the location of pixel $m$, then we write $d(\+p, \+q) = \+d_n,m$. Putting these components together yields the model in Figure~\ref{fig:model}. The generative story is:

\begin{itemize}
\item $\+\pi \sim \mbox{DirichletProcess}(\alpha)$
\item $k | \+\pi \sim \mbox{Discrete}(\+\pi)$
\item If $k$ is a new (empty) cluster:
  \begin{itemize}
  \item $\Lambda_k | T, \nu \sim \mbox{Wishart}(\Lambda | T, \nu)$ (in the notation of \cite{Murphy2003})
  \item $\+\mu_k | \kappa , \Lambda_k \sim \mathcal{N}\left( \+\mu_k | \+\mu_0 , (\kappa\Lambda)^{-1} \right)$
  \item $\beta_k | \alpha_0, \beta_0 \sim \mbox{Gamma}(\beta_k | \alpha_0, \mbox{rate} = \beta_0)$
  \end{itemize}
\item $\+x | \+\mu_k , \Lambda_k, \beta_k \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \mbox{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \mbox{rate} = \beta_k \right)$ (in the notation of section~\ref{sec:geodesic})
\end{itemize}

@@

\subsection{Geodesic Distance}
\label{sec:geodesic}
Figure~\ref{fig:radon_bd} illustrates that cells need not be elliptical, or even convex. All that can be said is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response. Therefore, shape constraints based on Euclidean distances are not justified. Instead, we want to encode the a priori belief that "the cell body should not spill outside its membrane." We quantify this property with a \emph{geodesic distance}. Construct a graph whose vertices are the pixels and undirected edge for adjacent pixels ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$). Each edge has a weight $w(i,j) = \left| b_i - b_j \right|$ where $b_k$ is the boundary Radon-like response at pixel $k$. We define $d(\+{x}_i, \+{x}_j)$ to be the total weight of the shortest path between pixels $i$ and $j$, which is computed with Johnson's algorithm. In practice, the $O(N^2)$ storage is prohibitive, so we instead we downsample the image into SLIC superpixels \cite{Achanta2012} and compute superpixel-based distances.Address practical concerns. Further, because individual cells are tiny relative to an entire slice, only local distances are needed, because far points must be outside the current cell. All paths crossing a cell boundary must traverse pixels with low, high, and again low boundary feature responses. Thus, the shortest path wil also be one with a high weight. So if pixels $i$ and $j$ do not belong to the same cell, defined by the region enclosed by a loop of pixels with high boundary feature response, then $d(\+{x}_i, \+{x}_j)$ is high.

@@

\section{Inference}
Our model encodes spatial locality in two novel ways: in the Gaussian covariance which bias class membership to Euclidean-close points, and in the geodesic distances which penalizes boundary crossings.

\section{Results}
Forthcoming.
